Constraints on quasi--normal-mode frequencies with LIGO-Virgo binary--black-hole observations

### 2021-04-01 papers/intro_paper/intro_paper.tex
removed colour from responses; cleaned remaining comments from tex

%\documentclass[twocolumn,prd,superscriptaddress,amsfonts,amssymb,amsmath,preprintnumbers,nofootinbib]{revtex4-1}
\documentclass[twocolumn,prd,aps,superscriptaddress,preprintnumbers,tightenlines,showpacs,nofootinbib,eqsecnum,amsfonts,amsmath]{revtex4-1}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[dvipsnames]{xcolor}
\usepackage{bm}
\usepackage{times}
\usepackage{xspace}
\usepackage[varg]{txfonts}
\usepackage[normalem]{ulem} % To get strikethrough (\sout)
\usepackage[colorlinks]{hyperref}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{url}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage{lineno}
%\linenumbers
\usepackage{makecell}

@@

\newcommand{\paperone}{Paper~I\xspace}
\newcommand{\abhi}[1]{\textcolor{Emerald}{#1}}
\newcommand{\rb}[1]{\textcolor{blue}{#1}}
\newcommand{\ab}[1]{\textcolor{cyan}{#1}}
\newcommand{\comment}[1]{\textcolor{red}{[#1]}}
\newcommand{\macro}[1]{\textcolor{orange}{\textbf{#1}}}

@@

%\title{Black hole spectroscopy using a complete gravitational wave signal from a binary black coalescence}

\title{Constraints on quasi--normal-mode frequencies with LIGO-Virgo binary--black-hole observations} 
%by making full use of gravitational-wave modeling}

@@

The consistency between the post-merger signal and the least damped QNM was first demonstrated in
Ref.~\cite{TheLIGOScientific:2016src} for GW150914, and confirmed through several independent analyses~\cite{Brito:2018rfr,Carullo:2019flw,Isi:2019aib,CalderonBustillo:2020tjf}. This was later extended to include
overtones in Refs.~\cite{Giesler:2019uxc,Isi:2019aib,Abbott:2020jks}. Consistency of the late-time waveform with a single QNM is a test of the ringdown of
a BBH coalescence, but not necessarily a test of the no-hair
conjecture, which requires the measurement of (at least) two QNMs (BH
spectroscopy), and consistency between them~\cite{Dreyer:2003bv,Gossan:2011ha, Meidam:2014jpa,Carullo:2018sfu,Carullo:2018gah,Isi:2019aib,Bhagwat:2019bwv,Cabero:2019zyt,Maselli:2019mjd}. The nature of the remnant object has also been explored through tests of BH thermodynamics, including the Hawking's area
theorem~\cite{Cabero:2017avf,Isi:2020tac}, the Bekenstein-Hod universal bound~\cite{Carullo:2021yxh}, the BH area quantization~\cite{Foit:2016uxn,Laghi:2020rgl}, \rb{the consistency of the merger frequency with predictions from NR simulations~\cite{Carullo:2018gah}} or through search for echos in the post-merger signal~\cite{Nielsen:2018lkf,Tsang:2019zra,Lo:2018sep,Abedi:2018npz,Abedi:2020sgg,Testa:2018bzd}.  None of these tests have found evidence for non-BH nature of the remnant
object (as described in GR) in LIGO-Virgo BBH observations.

Ground-based detectors, such as LIGO and Virgo, are most sensitive to
stellar-mass BH binaries \rb{that form remnants with mass $\gtrsim 50 M_{\odot}$ and} merge near the
minimum of their sensitivity band ($\sim 100$ Hz). As a consequence, the
remnant object \textit{rings down} at frequencies where the sensitivity is worse, leaving
very little signal-to-noise ratio (SNR) in the
post-merger signal. Most of the tests mentioned above restrict the data to this post-merger signal \rb{which, while being agnostic to deviations in the pre-merger dynamics, significantly} reduces parameter-estimation capabilities. Added to a lack of SNR is the ambiguity in clearly defining a start time for the ringdown phase (discussed, for example, in Refs.~\cite{Berti:2007fi,Baibhav:2017jhs,Bhagwat:2017tkm}). In some tests~\cite{Carullo:2018gah,Carullo:2019flw}, the ringdown start time has been left as a
free parameter to be estimated directly from the data. Whereas, in other cases, the ringdown start time is predicted from corresponding full-signal parameter-estimation analyses (see \texttt{pyRing} analysis in Ref.~\cite{Abbott:2020jks}). Uncertainties
in estimates of the ringdown start-time, as well as an overall lack of
SNR in the post-merger signal, given typical sensitivities of
ground-based detectors, \rb{can result in significant statistical uncertainties in the measurement of some of the QNM properties. This is especially true for the measurement of the QNM damping times which are, in general, harder to measure from the data}. Hence, one might want to look at alternate methods to measure QNMs using information from as much as the signal as possible.

An independent approach to BH spectroscopy, based on the full-signal
analysis, was introduced in Ref.~\cite{Brito:2018rfr}
(henceforth referred to as \paperone). Unlike methods that focus only
on the post-merger signal, it employs the
complete inspiral-merger-ringdown
(IMR) waveform to measure the complex QNM
frequencies. \ab{In the current version, this method assumes that GR provides a very good description of the signal up to merger, while being agnostic about the complex QNM frequencies. This allows to access the full SNR of the signal, reducing
measurement uncertainties. However, we note that this method can incorporate deviations from GR, notably from the (known) PN coefficients,  
also during the long inspiral stage~\cite{Abbott:2018lct,
LIGOScientific:2019fpa,Abbott:2020jks}, and can in principle incorporate deviations from GR, notably from the parameters calibrated to NR simulations,  
in the late inspiral and plunge.} Moreover, in this method, the definition of the
ringdown start time is built into the merger-ringdown model and does not need to be
either left as an additional free parameter or fixed using alternate
definitions. While \paperone presented the method and tested it for non-spinning BBHs, here we extend the analysis to the more
realistic astrophysical case in which BHs carry spin. Furthermore,
the IMR waveform model used in this paper is more accurate than what employed in \paperone,
because it contains higher-order corrections in PN theory and it was calibrated to a much larger set
of numerical-relativity (NR) waveforms~\cite{Bohe:2016gbl}. All astrophysical BHs are expected to be spinning,
and ignoring effects of spin has been shown to introduce systematic biases in the measurement of the
source properties.

@@

To demonstrate the robustness of the method in detecting possible
deviations from GR, we inject synthetic GW signals which are identical
to the corresponding GR prediction up to merger, and differ in their
post-merger description. We again choose binary-parameters similar to
GW150914 and GW190521 (see Table ~\ref{tab:injection_values}), but set
$\df{220} = \dtau{220} = 0.1 $.  In other words, we assume that the
frequency and damping time of our non-GR signal is 10\% more than the
corresponding GR prediction, although the pre-merger signal is
identical to GR. In Fig.~\ref{fig:nongr_waveform} we show this non-GR
waveform, \texttt{pSEOBNR} with respect to the original GR template,
\texttt{SEOBNR}. We see that the waveforms are identical in amplitude
and instanteneous frequency upto the merger (lower panel) , beyond
which the red (GR template) and blue (non-GR template) differ. We
summarize the results of the Bayesian analysis in
Fig.~\ref{fig:simulated_signal_nonGR} where we show the posterior
probability distributions for $(\df{220}, \dtau{220})$, or
equivalently $(\fngr{220}, \taungr{220})$. We find that they are
consistent with the corresponding values of the injection parameters,
indicated by the black dashed lines.  \abhi{We also note that at the
  SNRs we consider, since statistical uncertainties dominate
  systematic biases, the measurement excludes, at the 90\% credible
  level, the GR prediction \ab{of} the frequency ($\df{220}=0$), but
  includes the prediction \ab{of} the damping time
  ($\dtau{220}=0$). However, with louder events, one would expect
  these measurement errors to shrink and the $\taungr{220}$
  measurement to be inconsistent with the GR prediction, as well.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% GW150914-like signal with different deviations: GR vs nonGR recovery
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}%[h!]
        \includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_gr_m1m2.png}\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_ngr_m1m2.png}
        \includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_gr_a1za2z.png}\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_ngr_a1za2z.png}
        \includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_gr_fgrtaugr.png}\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_ngr_fgrtaugr.png}
        \includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_gr_Mfaf.png}\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_ngr_Mfaf.png}
         \caption{Comparison of a GW150914-like injection's parameters when signals with zero, 10\% and 50\% deviations ($\df{220} = \dtau{220} = 0, 0.1, 0.5$) are recovered using a GR ($\SEOB$) (left panels) or a non-GR ($\pSEOB$) (right panels) waveform model. The panels (from top to bottom) show the 2D posteriors (with 90\% credible levels) and corresponding marginalised 1D posteriors (with 90\% credible levels) in (detector-frame) masses (first row), dimensionless spins (second row), GR predictions of frequency and damping time (third row) and the remnant mass and spin predictions ($M_f$, $\chi_f$) \abhi{(fourth row)} from the frequency and damping time. In each case, the injection values are indicated by the black dashed lines. In the lowermost panel, the injection values of the final mass and spin, correspond to the injection with no deviations. The $\df{220} = \dtau{220} = 0.1$ signal is identical to the results shown in Fig.~\ref{fig:simulated_signal_nonGR}.}
        \label{fig:gr_ngr_comparison}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We additionally investigate the effects of erroneously assuming that
an underlying non-GR signal can be well-described by a GR one. We do
this by estimating the parameters of our non-GR signals using the GR
waveform model $\SEOB$ instead of the parameterized $\pSEOB$.  The
resulting one- and two-dimensional posteriors are shown in the right
panels of Fig.~\ref{fig:simulated_signal_nonGR} by purple curves for the
GW150914-like (top) and GW190521-like (bottom) signals
respectively. For both signals, we find the \abhi{2D} $\SEOB$ estimates are
markedly biased with respect to the $\pSEOB$ estimates.  We investigate 
this impact on the estimation of the GR parameters $\bxigr$ by changing the
magnitude of deviation in the non-GR parameters more in Fig.~\ref{fig:gr_ngr_comparison}. We restrict ourselves 
to a GW150914-like event, and for comparison, add a synthetic signal
with $\df{220} = \dtau{220} = 0.5$, or 50 \% deviation, alongside the
10\% non-GR signal and a signal with no deviation (essentially GR)
mentioned above. The left (right) panels show the
posterior probability distributions of the three signals using the GR,
$\SEOB$ (non-GR, $\pSEOB$) waveform model. The first difference we
note is the $\SEOB$ recoveries yield biased estimates when the
underlying signal is non-GR, while the $\pSEOB$ recoveries do
not. Furthermore, as we increase the deviations, while the $\SEOB$
recoveries expectedly get more biased, the $\pSEOB$ measurements are
robust in consistently recovering the injected value. This gives us 
confidence that the $\pSEOB$ model can accurately measure QNM 
properties without biasing measurements of inspiral quantities, like masses and spins.
\ab{To compute the samples of the final mass and spin in the last row, we start from the samples
of the complex QNM frequencies $(\fngr{220}, \taungr{220})$, which are obtained from the 
fractional deviation samples $(\df{220},\dtau{220})$ and the GR-quantities
$(\fgr{220},\taugr{220})$ using Eqs.~(\ref{eq:nongr_freqs_a}) and Eq.~(\ref{eq:nongr_freqs_b})},
\abhi{and then invert them \sout{$(\fngr{220}, \taungr{220})$} using the fitting formula in Ref.~\cite{Berti:2005ys}.
The three BBH signals, $\df{220} = \dtau{220} = 0, 0.1, 0.5$, correspond to
three unique sets of values for $(M_f,\chi_f)$. These predicted values are correctly 
recovered by the $\pSEOB$ waveform (bottom-right plot in Fig.~\ref{fig:gr_ngr_comparison}),
leading to three distinct and disjointed 2D posteriors on $(M_f,\chi_f)$, unlike the $\SEOB$
analysis (to reduce clutter in the plot, we just plot the injection values for $\df{220} = \dtau{220} = 0$. }

@@

%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
\label{sec:acknowledgements}
The authors thank Max Isi for generously allowing us to use the code developed in Ref.~\cite{Isi:2019asy} for the hierarchical inference analysis, and Gregorio Carullo for carefully reading the manuscript and providing useful comments. The authors would like to thank everyone at the frontline of the Covid-19 pandemic. \abhi{The authors are grateful for computational resources provided by the LIGO Laboratory and supported by the National Science Foundation Grants PHY-0757058 and PHY-0823459, as well as computational resources at the AEI, specifically the Hypatia cluster.} R.B. acknowledges financial support from the European Union's Horizon 2020 research and innovation programme under the Marie Sk\l odowska-Curie grant agreement No. 792862, from the European Union's H2020 ERC, Starting Grant agreement no.~DarkGRA--757480, and from the MIUR PRIN and FARE programmes (GW-NEXT, CUP:~B84I20000100001). We also acknowledge support from the Amaldi Research Center funded by the MIUR program ``Dipartimento di Eccellenza'' (CUP:~B81I18001170001).
%%%%%%%%%%%%%%%%%

@@

We choose a \abhi{spinning}, precessing NR-surrogate model \texttt{NRSur}~\footnote{This waveform model is 
called \texttt{NRSur7dq4} in LAL.} (valid up to mass ratio 4) to simulate the actual GW190521 signal
observed by the LIGO and Virgo detectors~\cite{Abbott:2020tfl} (see
Table I of Ref.~\cite{Abbott:2020tfl})). The choice of the
\texttt{NRSur} model is motivated by the fact that it is the most
accurate model in the parameter range described by GW190521, because 
it is built by directly interpolating NR waveforms. In
Fig.~\ref{fig:21g_systematics}, we indicate with a black cross what
the injected \texttt{NRSur} signal predicts for the QNM $(\ell=2,m=2)$
frequency and damping time. For comparison, we also show with a
black solid curve the results obtained when recovering the actual
signal GW190521 with the $\pSEOB$ model. As seen in the plot,
while the measurement of the frequency is consistent with the
prediction, we overestimate the damping time.

>>>>

\documentclass[twocolumn,prd,aps,superscriptaddress,preprintnumbers,tightenlines,showpacs,nofootinbib,eqsecnum,amsfonts,amsmath]{revtex4-1}
\usepackage{epsfig}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{bm}
\usepackage[dvipsnames]{xcolor}
\usepackage{bm}
\usepackage{times}
\usepackage{xspace}
\usepackage[varg]{txfonts}
\usepackage[normalem]{ulem} % To get strikethrough (\sout)
\usepackage[colorlinks]{hyperref}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{url}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage{lineno}
%\linenumbers
\usepackage{makecell}

@@

\newcommand{\paperone}{Paper~I\xspace}

@@

\title{Constraints on quasi--normal-mode frequencies with LIGO-Virgo binary--black-hole observations}

@@

The consistency between the post-merger signal and the least damped QNM was first demonstrated in
Ref.~\cite{TheLIGOScientific:2016src} for GW150914, and confirmed through several independent analyses~\cite{Brito:2018rfr,Carullo:2019flw,Isi:2019aib,CalderonBustillo:2020tjf}. This was later extended to include
overtones in Refs.~\cite{Giesler:2019uxc,Isi:2019aib,Abbott:2020jks}. Consistency of the late-time waveform with a single QNM is a test of the ringdown of
a BBH coalescence, but not necessarily a test of the no-hair
conjecture, which requires the measurement of (at least) two QNMs (BH
spectroscopy), and consistency between them~\cite{Dreyer:2003bv,Gossan:2011ha, Meidam:2014jpa,Carullo:2018sfu,Carullo:2018gah,Isi:2019aib,Bhagwat:2019bwv,Cabero:2019zyt,Maselli:2019mjd}. The nature of the remnant object has also been explored through tests of BH thermodynamics, including the Hawking's area
theorem~\cite{Cabero:2017avf,Isi:2020tac}, the Bekenstein-Hod universal bound~\cite{Carullo:2021yxh}, the BH area quantization~\cite{Foit:2016uxn,Laghi:2020rgl}, the consistency of the merger frequency with predictions from NR simulations~\cite{Carullo:2018gah} or through search for echos in the post-merger signal~\cite{Nielsen:2018lkf,Tsang:2019zra,Lo:2018sep,Abedi:2018npz,Abedi:2020sgg,Testa:2018bzd}.  None of these tests have found evidence for non-BH nature of the remnant
object (as described in GR) in LIGO-Virgo BBH observations.

Ground-based detectors, such as LIGO and Virgo, are most sensitive to
stellar-mass BH binaries that form remnants with mass $\gtrsim 50 M_{\odot}$ and merge near the
minimum of their sensitivity band ($\sim 100$ Hz). As a consequence, the
remnant object \textit{rings down} at frequencies where the sensitivity is worse, leaving
very little signal-to-noise ratio (SNR) in the
post-merger signal. Most of the tests mentioned above restrict the data to this post-merger signal which, while being agnostic to deviations in the pre-merger dynamics, significantly reduces parameter-estimation capabilities. Added to a lack of SNR is the ambiguity in clearly defining a start time for the ringdown phase (discussed, for example, in Refs.~\cite{Berti:2007fi,Baibhav:2017jhs,Bhagwat:2017tkm}). In some tests~\cite{Carullo:2018gah,Carullo:2019flw}, the ringdown start time has been left as a
free parameter to be estimated directly from the data. Whereas, in other cases, the ringdown start time is predicted from corresponding full-signal parameter-estimation analyses (see \texttt{pyRing} analysis in Ref.~\cite{Abbott:2020jks}). Uncertainties
in estimates of the ringdown start-time, as well as an overall lack of
SNR in the post-merger signal, given typical sensitivities of
ground-based detectors, can result in significant statistical uncertainties in the measurement of some of the QNM properties. This is especially true for the measurement of the QNM damping times which are, in general, harder to measure from the data. Hence, one might want to look at alternate methods to measure QNMs using information from as much as the signal as possible.

An independent approach to BH spectroscopy, based on the full-signal
analysis, was introduced in Ref.~\cite{Brito:2018rfr}
(henceforth referred to as \paperone). Unlike methods that focus only
on the post-merger signal, it employs the
complete inspiral-merger-ringdown
(IMR) waveform to measure the complex QNM
frequencies. In the current version, this method assumes that GR provides a very good description of the signal up to merger, while being agnostic about the complex QNM frequencies. This allows to access the full SNR of the signal, reducing
measurement uncertainties. However, we note that this method can incorporate deviations from GR, notably from the (known) PN coefficients,  
also during the long inspiral stage~\cite{Abbott:2018lct,
LIGOScientific:2019fpa,Abbott:2020jks}, and can in principle incorporate deviations from GR, notably from the parameters calibrated to NR simulations,  
in the late inspiral and plunge. Moreover, in this method, the definition of the
ringdown start time is built into the merger-ringdown model and does not need to be
either left as an additional free parameter or fixed using alternate
definitions. While \paperone presented the method and tested it for non-spinning BBHs, here we extend the analysis to the more
realistic astrophysical case in which BHs carry spin. Furthermore,
the IMR waveform model used in this paper is more accurate than what employed in \paperone,
because it contains higher-order corrections in PN theory and it was calibrated to a much larger set
of numerical-relativity (NR) waveforms~\cite{Bohe:2016gbl}. All astrophysical BHs are expected to be spinning,
and ignoring effects of spin has been shown to introduce systematic biases in the measurement of the
source properties.

@@

To demonstrate the robustness of the method in detecting possible
deviations from GR, we inject synthetic GW signals which are identical
to the corresponding GR prediction up to merger, and differ in their
post-merger description. We again choose binary-parameters similar to
GW150914 and GW190521 (see Table ~\ref{tab:injection_values}), but set
$\df{220} = \dtau{220} = 0.1 $.  In other words, we assume that the
frequency and damping time of our non-GR signal is 10\% more than the
corresponding GR prediction, although the pre-merger signal is
identical to GR. In Fig.~\ref{fig:nongr_waveform} we show this non-GR
waveform, \texttt{pSEOBNR} with respect to the original GR template,
\texttt{SEOBNR}. We see that the waveforms are identical in amplitude
and instanteneous frequency upto the merger (lower panel) , beyond
which the red (GR template) and blue (non-GR template) differ. We
summarize the results of the Bayesian analysis in
Fig.~\ref{fig:simulated_signal_nonGR} where we show the posterior
probability distributions for $(\df{220}, \dtau{220})$, or
equivalently $(\fngr{220}, \taungr{220})$. We find that they are
consistent with the corresponding values of the injection parameters,
indicated by the black dashed lines.  We also note that at the
  SNRs we consider, since statistical uncertainties dominate
  systematic biases, the measurement excludes, at the 90\% credible
  level, the GR prediction of the frequency ($\df{220}=0$), but
  includes the prediction of the damping time
  ($\dtau{220}=0$). However, with louder events, one would expect
  these measurement errors to shrink and the $\taungr{220}$
  measurement to be inconsistent with the GR prediction, as well.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% GW150914-like signal with different deviations: GR vs nonGR recovery
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}%[h!]
        \includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_gr_m1m2.png}\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_ngr_m1m2.png}
        \includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_gr_a1za2z.png}\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_ngr_a1za2z.png}
        \includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_gr_fgrtaugr.png}\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_ngr_fgrtaugr.png}
        \includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_gr_Mfaf.png}\includegraphics[width=0.5\textwidth]{figures/GW150914_simulated_signal_0p0_0p1_0p5_ngr_Mfaf.png}
         \caption{Comparison of a GW150914-like injection's parameters when signals with zero, 10\% and 50\% deviations ($\df{220} = \dtau{220} = 0, 0.1, 0.5$) are recovered using a GR ($\SEOB$) (left panels) or a non-GR ($\pSEOB$) (right panels) waveform model. The panels (from top to bottom) show the 2D posteriors (with 90\% credible levels) and corresponding marginalised 1D posteriors (with 90\% credible levels) in (detector-frame) masses (first row), dimensionless spins (second row), GR predictions of frequency and damping time (third row) and the remnant mass and spin predictions ($M_f$, $\chi_f$) (fourth row) from the frequency and damping time. In each case, the injection values are indicated by the black dashed lines. In the lowermost panel, the injection values of the final mass and spin, correspond to the injection with no deviations. The $\df{220} = \dtau{220} = 0.1$ signal is identical to the results shown in Fig.~\ref{fig:simulated_signal_nonGR}.}
        \label{fig:gr_ngr_comparison}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We additionally investigate the effects of erroneously assuming that
an underlying non-GR signal can be well-described by a GR one. We do
this by estimating the parameters of our non-GR signals using the GR
waveform model $\SEOB$ instead of the parameterized $\pSEOB$.  The
resulting one- and two-dimensional posteriors are shown in the right
panels of Fig.~\ref{fig:simulated_signal_nonGR} by purple curves for the
GW150914-like (top) and GW190521-like (bottom) signals
respectively. For both signals, we find the 2D $\SEOB$ estimates are
markedly biased with respect to the $\pSEOB$ estimates.  We investigate 
this impact on the estimation of the GR parameters $\bxigr$ by changing the
magnitude of deviation in the non-GR parameters more in Fig.~\ref{fig:gr_ngr_comparison}. We restrict ourselves 
to a GW150914-like event, and for comparison, add a synthetic signal
with $\df{220} = \dtau{220} = 0.5$, or 50 \% deviation, alongside the
10\% non-GR signal and a signal with no deviation (essentially GR)
mentioned above. The left (right) panels show the
posterior probability distributions of the three signals using the GR,
$\SEOB$ (non-GR, $\pSEOB$) waveform model. The first difference we
note is the $\SEOB$ recoveries yield biased estimates when the
underlying signal is non-GR, while the $\pSEOB$ recoveries do
not. Furthermore, as we increase the deviations, while the $\SEOB$
recoveries expectedly get more biased, the $\pSEOB$ measurements are
robust in consistently recovering the injected value. This gives us 
confidence that the $\pSEOB$ model can accurately measure QNM 
properties without biasing measurements of inspiral quantities, like masses and spins.
To compute the samples of the final mass and spin in the last row, we start from the samples
of the complex QNM frequencies $(\fngr{220}, \taungr{220})$, which are obtained from the 
fractional deviation samples $(\df{220},\dtau{220})$ and the GR-quantities
$(\fgr{220},\taugr{220})$ using Eqs.~(\ref{eq:nongr_freqs_a}) and Eq.~(\ref{eq:nongr_freqs_b}),
and then invert them using the fitting formula in Ref.~\cite{Berti:2005ys}.
The three BBH signals, $\df{220} = \dtau{220} = 0, 0.1, 0.5$, correspond to
three unique sets of values for $(M_f,\chi_f)$. These predicted values are correctly 
recovered by the $\pSEOB$ waveform (bottom-right plot in Fig.~\ref{fig:gr_ngr_comparison}),
leading to three distinct and disjointed 2D posteriors on $(M_f,\chi_f)$, unlike the $\SEOB$
analysis (to reduce clutter in the plot, we just plot the injection values for $\df{220} = \dtau{220} = 0$.

@@

%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
\label{sec:acknowledgements}
The authors thank Max Isi for generously allowing us to use the code developed in Ref.~\cite{Isi:2019asy} for the hierarchical inference analysis, and Gregorio Carullo for carefully reading the manuscript and providing useful comments. The authors would like to thank everyone at the frontline of the Covid-19 pandemic. The authors are grateful for computational resources provided by the LIGO Laboratory and supported by the National Science Foundation Grants PHY-0757058 and PHY-0823459, as well as computational resources at the AEI, specifically the Hypatia cluster. R.B. acknowledges financial support from the European Union's Horizon 2020 research and innovation programme under the Marie Sk\l odowska-Curie grant agreement No. 792862, from the European Union's H2020 ERC, Starting Grant agreement no.~DarkGRA--757480, and from the MIUR PRIN and FARE programmes (GW-NEXT, CUP:~B84I20000100001). We also acknowledge support from the Amaldi Research Center funded by the MIUR program ``Dipartimento di Eccellenza'' (CUP:~B81I18001170001).
%%%%%%%%%%%%%%%%%

@@

We choose a spinning, precessing NR-surrogate model \texttt{NRSur}~\footnote{This waveform model is 
called \texttt{NRSur7dq4} in LAL.} (valid up to mass ratio 4) to simulate the actual GW190521 signal
observed by the LIGO and Virgo detectors~\cite{Abbott:2020tfl} (see
Table I of Ref.~\cite{Abbott:2020tfl})). The choice of the
\texttt{NRSur} model is motivated by the fact that it is the most
accurate model in the parameter range described by GW190521, because 
it is built by directly interpolating NR waveforms. In
Fig.~\ref{fig:21g_systematics}, we indicate with a black cross what
the injected \texttt{NRSur} signal predicts for the QNM $(\ell=2,m=2)$
frequency and damping time. For comparison, we also show with a
black solid curve the results obtained when recovering the actual
signal GW190521 with the $\pSEOB$ model. As seen in the plot,
while the measurement of the frequency is consistent with the
prediction, we overestimate the damping time.
